---
title: "Chocolate bar ratings"
author: "Lee Olney"
date: "2022-11-15"
format: 
  html:
    toc: true
    code-fold: true
execute: 
  warning: false
editor: visual
---

This Quarto document contains a table built in R using [{gt}](https://gt.rstudio.com/) and [{gtExtras}](https://jthomasmock.github.io/gtExtras/) packages for the [R Studio 2022 Table Contest](https://www.rstudio.com/blog/rstudio-table-contest-2022/) submission. The table was created for [#TidyTuesday](https://github.com/rfordatascience/tidytuesday) week 3 [Chocolate Bar ratings](https://github.com/rfordatascience/tidytuesday/blob/master/data/2022/2022-01-18/readme.md), [shared](https://twitter.com/leeolney3/status/1483378919957008387) on Twitter and Github on Jan 18, 2022.

The table summarizes plain dark chocolate bars' ratings (between 1 to 5) by twenty two manufacturers with more than fifteen reviews, data from [Flavors of Cacao](http://flavorsofcacao.com/chocolate_database.html) credits to [Georgios Karamanis and Kelsey E Gonzalez](https://github.com/rfordatascience/tidytuesday/issues/408). The table uses an inline box plot function by [Thomas Mock](https://themockup.blog/about.html) in the blog post [Embedding custom HTML in gt tables](https://themockup.blog/posts/2020-10-31-embedding-custom-features-in-gt-tables/).

```{r load libraries}
#| message: false
library(tidyverse)
library(gt)
library(gtExtras)
library(kableExtra)
```

```{r prepare data}
#| message: false
choco<- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-18/chocolate.csv')
```

```{r table data}
# Get company_manufacturer with more than 15 reviews
df1 = choco %>% count(company_manufacturer, sort=T) %>% filter(n>15)

# Table data
df2 = choco %>% 
  filter(company_manufacturer %in% df1$company_manufacturer)  %>% 
  group_by(company_manufacturer) %>%
  mutate(avg=mean(rating)) %>%
  ungroup() %>%
  mutate(rank = dense_rank(desc(avg))) %>%
  arrange(rank)

rate_list <- split(df2$rating, df2$rank)
rate_rng <- range(df2$rating)
```

```{r inline boxplot function}
# Citation: Mock (2020, Oct. 31). The Mockup Blog: Embedding custom HTML in gt tables. Retrieved from https://themockup.blog/posts/2020-10-31-embedding-custom-features-in-gt-tables/
gt_plot <- function(table_data, column, plot_data, plot_fun, ...){
  text_transform(
    table_data,
    # note the use of {{}} here - this is tidy eval
    # that allows you to indicate specific columns
    locations = cells_body(columns = vars({{column}})),
    fn = function(x){
      plot <- map(plot_data, plot_fun, width = 300, height = 70, same_lim = TRUE, ...)
      plot_svg <- map(plot, "svg_text")
      map(plot_svg, gt::html)
    }
  )
}
```

```{r table}
#| warning: false
# replaced gt_sparkline() with gt_plt_dist() in gtExtras version 0.4.4.9000
# replaced colors with palette in gt_merge_stack in gtExtras version 0.4.4.9000
df2 %>%
  mutate(company_manufacturer==case_when(company_manufacturer=="Smooth Chocolator, The" ~ "The Smooth Chocolator", 
                                         TRUE~company_manufacturer)) %>%
  group_by(company_location,company_manufacturer) %>%
  summarise(n=n(),
            average = round(mean(rating),2),
            min=min(rating),
            median = round(median(rating),2),
            max=max(rating),
            range= max-min,
            histogram=list(rating),
            .groups="drop") %>%
  arrange(desc(average)) %>%
  mutate(boxplot ="",
         n2 = n) %>%
  select(company_location, company_manufacturer, n, n2, average, histogram, min, median, max, range, boxplot) %>%
  gt() %>%
  gt_theme_538() %>%
  gt_plt_dist(histogram, 
              type = "histogram", 
              line_color = "#66462c", 
              fill_color = "#66462c", 
              bw = .25, 
              same_limit = TRUE) %>%
  gt_plot(
    column = boxplot,  # column to create plot in 
    plot_data = rate_list, # external data to reference
    plot_fun = spec_boxplot,  # which plot fun
    lim = rate_rng, # range applied
    ) %>%
  gt_plt_bar(column=n2, 
             color="#82a6b1",
             width=30) %>%
  gt_merge_stack(company_manufacturer, company_location, palette=c("#38160d","grey")) %>%
  gt_color_rows(columns = c("average","range"),
                palette = "ggsci::brown_material") %>%
  cols_align(columns = c("histogram", "boxplot", "median"),
             align="center") %>%
  cols_label(company_manufacturer = html("Manufacturer"),
             n=html(""),
             n2=html("N reviewed")) %>%
  tab_spanner(label="Rating", 
              columns=c(average:boxplot)) %>%
  tab_header(title=md("<span style='color:#411d13'>Ratings of Plain Dark Chocolate Bar</span>"),
             subtitle=md("Summary table of ratings (between 1 to 5) by 22 manufacturers with more than 15 reviews, according to *Flavors of Cacao*.")) %>%
  tab_source_note(source_note = gt::html("<br>#TidyTuesday Week 3  |  Data source: Flavors of Cacao, by way of Georgios and Kelsey  |  Inline boxplot function from Thomas Mock")) %>%
  # Adjust sub-title font
  tab_style(
    style = list(
      cell_text(
        weight="lighter"
      )
    ),
    locations = list(
      cells_title(groups = "subtitle")
    )
  )  
```
