---
title: "NYTimes best sellers"
author: "Lee Olney"
date: "2022-11-15"
format: 
  html:
    toc: true
    code-fold: true
execute: 
  warning: false
editor: visual
---

This Quarto document contains a table built in R using [{gt}](https://gt.rstudio.com/) and [{gtExtras}](https://jthomasmock.github.io/gtExtras/) packages for the [R Studio 2022 Table Contest](https://www.rstudio.com/blog/rstudio-table-contest-2022/) submission.

The table was created for [#TidyTuesday](https://github.com/rfordatascience/tidytuesday) week 19 [NYTimes best sellers](https://github.com/rfordatascience/tidytuesday/blob/master/data/2022/2022-05-10/readme.md), [shared](https://twitter.com/leeolney3/status/1523875516171833344) on Twitter and Github on May 10, 2022. The table shows 21 titles with 80 or more total weeks on The New York Times' fiction bestseller list between 1931 and 2020, arranged in descending order of total weeks appeared. Data from [Post45 Data](http://flavorsofcacao.com/chocolate_database.html), credits to [Sara Stoudt](https://github.com/rfordatascience/tidytuesday/issues/434).

```{r load libraries}
#| message: false
library(tidyverse)
library(gt)
library(gtExtras)
```

```{r import data}
#| message: false
nyt_titles <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-05-10/nyt_titles.tsv')
nyt_full <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-05-10/nyt_full.tsv')
```

```{r prepare data}
df1 = nyt_titles %>% slice_max(total_weeks, n=20) %>%
  mutate(title=str_to_title(title)) 

df2 = nyt_full %>% filter(title_id %in% df1$id) %>%
  mutate(rank=-1*rank) %>%
  group_by(id=title_id) %>%
  arrange(week) %>%
  summarise(timeline=list(rank),.groups="drop")

df3 = df1 %>% left_join(df2, by="id")
```

```{r table}
# replaced gt_sparkline() with gt_plt_sparkline() in gtExtras version 0.4.4.9000
# amended title
df3 %>%
  select(-id) %>%
  mutate(wk = total_weeks) %>%
  relocate(wk,.after=total_weeks) %>%
  gt() %>%
  gt_theme_nytimes() %>%
  #gt_sparkline(timeline, label=F, range_colors = c("#7014f2", "#06b178"))
  gt_plt_sparkline(timeline, label=F,  palette = c("black", "black", "#7014f2", "#06b178", "lightgrey")) %>%
  gt_plt_bar_pct(wk, fill="#ffc300") %>%
  cols_width(wk~px(100),
             first_week~px(100)) %>%
  cols_align(align="right", columns=first_week) %>%
  cols_label(total_weeks="total weeks",
             first_week="first week",
             debut_rank="debut rank",
             best_rank="best rank",
             wk="") %>%
  tab_header(title="NY Times bestsellers",
             subtitle=md("List of titles with more than 80 total weeks on the ***The New York Times***' fiction bestseller list between 1931 and 2020, arranged in descending order of total weeks.")) %>%
  tab_source_note("#TidyTuesday week 19 | Data from Post45 Data by way of Sara Stoudt") %>%
  tab_style(style = list(cell_text(style = "italic", color="black")),
    locations = cells_body(columns = title))
```
